The \dealii Library, Version 9.4

### 2022-06-24 9.4/paper.tex
draft for sec 2.7 large MPI

   A new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages containing more than $2^{31}$ objects.
  This library either uses the new MPI-4 functions, such as \texttt{MPI\_Send\_c()}, or an internal implementation of large objects for
  compatibility with MPI-3.

>>>>

   Sending large messages over MPI or reading and writing large buffers using MPI I/O,
   especially when dealing with small datatypes like \texttt{MPI\_CHAR},
   can often require processing more than $2^{31}$ objects at once. Prior
   to the recent MPI-4 standard, which introduced a new set of \texttt{MPI\_*\_c()} functions, all MPI functions expected the ``count''
   to be a signed integer. This limits the number of objects to $2^{31}$,
   i.e., writing of up to 2 GB at once if the datatype is \texttt{MPI\_CHAR}.

   Several places in the library could run into the issue mentioned above,
   for example when producing large graphical output (per rank), large
   checkpointing (per rank), or when broadcasting large datasets for lookup tables.

   With this release we introduce a new module \texttt{Utilities::MPI::LargeCount} which enables sending and receiving MPI messages and I/O containing more than $2^{31}$ objects by implementing
   the before-mentioned \texttt{MPI\_*\_c()} functions, i.e., \texttt{MPI\_Bcast\_c()}, using MPI-3 features if necessary. The
   solution is based on custom datatypes as described in~\ref{hammond2014int_max}.

   This functionality is now used in all places in the library, where
   large counts might be necessary (\texttt{DataOut::write\_vtu\_in\_parallel()} for graphical output, checkpointing of \texttt{Triangulation} objects, etc.).
   %
   While implementing these changes we also worked on the following:
   \begin{itemize}
    \item Testing of large I/O with large chunks per MPI rank (>2 GB) and
    large total sizes (4 GB +). Several instances of 32 bit datatypes for
    offsets were changed to 64 bit to correctly support files larger
    than 4 GB (HDF5 output, VTU output, checkpointing).
    \item Performance testing of MPI I/O routines used for parallel VTU output with large performance improvements by switching from a shared
    file pointer to individual file pointers, see Table~\ref{tab:mpi-io}.
   \end{itemize}

\begin{table}
  \caption{\it Performance comparison of writing large VTU graphical output using MPI I/O on
  Frontera /scratch1/ with 16 file servers (``OSTs'') and a theoretical peak performance of 50 GB/s. ROMIO seems to enforce sequential writes when using \texttt{MPI\_File\_write\_ordered()}.}
  \label{tab:mpi-io}

  \begin{tabular}{|c|c|c|c|}
  \hline
Version       & MPI routine used            & 1 file, striping 16 & 16 files, no striping \\
\hline
old VTU code  & \texttt{MPI\_File\_write\_ordered()} & 3 GB/s              & 15 GB/s              \\
new in 9.4 & \texttt{MPI\_File\_write\_at\_all()} & 17 GB/s             & 21 GB/s
\\
\hline
\end{tabular}

  \end{table}
